Stack Safety as a Security Property

### 2020-06-15 main.tex
Cleaned up judgments and wrote very briefly about heap safety

    \[\frac{\mach_1 \stepstoobs{\obs} \mach_2 \hspace{.5cm} (\mach_1, \pol_1) \stepstopol \pol_2}
           {\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}}\]
\leo{Notation: Should $\stepstoobs{}$ be partial?}

@@

We define a few operations on traces generically.
The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:
  \[\frac{f\ t}
         {\mathit{split}(f,tT) = t, tT}\]
  \[\frac{\neg f\ t \hspace{.5cm} \mathit{split}(f,T) = T_1,T_2}
         {\mathit{split}(f,tT) = tT_1,T_2}\]

The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
of the first (if it has one) with the first of the second:
  \[\frac{}
         {T_1t \cdot T_2 = T_1 T_2}\]

We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
\[\machT \in \mathit{MOTrace} = \mathit{Trace}\ (\MACHS \times \OBSS)\]
\[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]
%
\[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \mach_1 \hookrightarrow \machT}
       {\mach_0 \hookrightarrow (\mach_0,\obs) \machT}\]
%
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

\sna{Lambdas in the split are a bit busy, not sure anything will look better.}
We will also wish to take a prefix of the trace from an initial state up through the first state where
some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
\(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):
  \[\frac{\mach \hookrightarrow \machT \hspace{.5cm} \mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
            = \machT_1,\machT_2}
         {\mach \hookrightarrow \machT_1 | f}\]

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\):
%
\leo{I think this would read better as not exists. Thoughts?}
\sna{Looks good.}
  \[\MPT \in \mathit{MPOTrace} = \mathit{Trace}\ (\MACHS \times \POLS \times \OBSS)\]
  \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]
  \[\frac{\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1}
         {\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}}\]
  \[\frac{\mach_0 \stepstoobs{\obs} \mach_1 \hspace{.5cm} \pol_0 \rightharpoonup \pol_1 \hspace{.5cm}
            \mpstate{\mach_1}{\pol_1} \hookrightarrow \machT}
         {\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT}\]

We also lift the prefix-up-to relation to such traces.
  \[\frac{\mpstatename \hookrightarrow \MPT \hspace{.5cm}
            \mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2}
         {\mpstatename \hookrightarrow \MPT_1 | f}\]

@@

\[\frac{} {\obsT \lesssim \obsT}\]

\[\frac{} {\tau \lesssim \obsT}\]

\[\frac{\obsT_1 = \tau \obsT_1' \hspace{.5cm} \obsT_1' \lesssim \obsT_2}
       {\obsT_1 \lesssim \obsT_2}\]

\[\frac{\obsT_2 = \tau \obsT_2' \hspace{.5cm} \obsT_1 \lesssim \obsT_2'}
       {\obsT_1 \lesssim \obsT_2}\]

\[\frac{\obsT_1 = w \obsT_1' \hspace{.5cm} \obsT_2 = w \obsT_2' \hspace{.5cm} \obsT_1' \lesssim \obsT_2'}
       {\obsT_1 \lesssim \obsT_2}\]

@@

% Fixed by Sean
%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
\[\frac{\neg \isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \lesssim \pi_o(\machT)}
           {\MPT \sim_P \machT}\]

\[\frac{\isinf(\MPT) \hspace{.5cm} \pi_o(\MPT) \simeq \pi_o(\machT)}
           {\MPT \sim_P \machT}\]

@@

 \[\mathit{Cof}(\mach,\funid,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and } \codemap(\component) = \codemap(\mach(\PCname)) \\
    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS
                          \text{ and } \codemap(\component) = \funid \not = \codemap(\mach(\PCname)) \\
    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
                          \text{ and } \component \leq \mach(\SP) - n \\
    (\HIGHSEC,\LOWINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
                         \text{ and } \component \geq \mach(\SP) \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

@@

  \leo{Think more about whether we vary mc or step mc}
  \sna{With code, it must be step mc}

@@

    Some forms of security rely on protected control flow. In a classic
    example, a system call that launches a missile might be gated behind code
    that checks a password. A piece of malicious code might jump into the middle
    of the sequence and execute the system call without the appropriate password,
    and still obey integrity if the malicious code never returns to its caller.
    Integrity here protects data, not privileges. But such a jump does violate
    eager stack confidentiality if the attacker jumps beyond its own function's
    instructions without performing a designated call or returning to an
    appropriate site.

@@

    Discuss how we would do it, referring to Arthur's paper.  Probably not very long or formal?

@@

    Return integrity captures the expected behavior of marked returns with
    regard to their call points. We begin by defining a coinductive relation
    \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1 \MPT_2\). Given a call map
    \(\callmap\) and return map \(\retmap\), \(\MPT\) is related to
    \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\) whose head is the first
    unmatched return in \(\MPT\). So if \(\MPT\) starts with a return,
    it is related to itself:
%
    \[\frac{\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \retmap(\mach(\PCname))}
           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
              \mpstate{\mach}{\pol}}\]
    \[\frac{\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \retmap(\mach(\PCname))}
           {\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
              \mpstate{\mach}{\pol}\MPT}\]
%
    If \(\MPT\) starts with a non-call, non-return state, its unmatched
    return must come later:
%
    \[\frac{\begin{split}
            & \callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'
            \end{split}}
           {\mathit{Unmatched}\ \callmap\ \retmap\ 
            \mpstate{\mach}{\pol}\MPT\ \MPT'}\]
%   
    And most intricately, if \(\MPT\) starts with a call, we must find the
    return that matches it, which will be the unmatched return once we step
    past the call. Then the first unmatched return beyond it is the true
    unmatched return.
    \[\frac{\begin{split}
            & \callmap(\mach(\PCname)) = n
              \hspace{.5cm} \neg \retmap(\mach(\PCname)) \\
            & \MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}} \\
            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}} \\
            & \MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}} \\
            & \mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}} \\
            \end{split}}
            {\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}}\]

>>>>

    \judgmenttwo{\(\mach_1 \stepstoobs{\obs} \mach_2\)}{\((\mach_1, \pol_1) \stepstopol \pol_2\)}
           {\(\mpstate{\mach_1}{\pol_1} \stepstoobs{\obs}_P \mpstate{\mach_2}{\pol_2}\)}

\leo{Notation: Should $\stepstoobs{}$ be partial?}

@@

We define a few operations on traces generically.
The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\) and a predicate on elements \(f\) and
separates \(T\) into a prefix and suffix at the first element on which \(f\) holds:

  \judgment{\(f\ t\)}
           {\(\mathit{split}(f,tT) = t, tT\)}

  \judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
              {\(\mathit{split}(f,tT) = tT_1,T_2\)}

The {\it join} operation \(T_1 \cdot T_2\) takes two traces and links them by replacing the last element
of the first (if it has one) with the first of the second:

  \judgment{}
           {\(T_1t \cdot T_2 = T_1 T_2\)}

We define the ``trace-of'' operator \(\hookrightarrow\) that coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
\[\machT \in \mathit{MOTrace} = \mathit{Trace}\ (\MACHS \times \OBSS)\]
\[\hookrightarrow : \machT \rightarrow \mathit{MOTrace}\]

\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}

Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

\sna{Lambdas in the split are a bit busy, not sure anything will look better.}
We will also wish to take a prefix of the trace from an initial state up through the first state where
some condition holds, using {\it split}. Let \(f\) be a predicate on machine states. Then we define
\(\mach \hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of \(\machT\) up to \(f\)''):

  \judgmenttwo{\(\mach \hookrightarrow \machT\)}
              {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT) = \machT_1,\machT_2\)}
              {\(\mach \hookrightarrow \machT_1 | f\)}

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\):

\leo{I think this would read better as not exists. Thoughts?}
\sna{Looks good.}
  \[\MPT \in \mathit{MPOTrace} = \mathit{Trace}\ (\MACHS \times \POLS \times \OBSS)\]
  \[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \mathit{MPOTrace}\]

  \judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
           {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}

  \judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
                {\(\pol_0 \rightharpoonup \pol_1\)}
                {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
      {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}

We also lift the prefix-up-to relation to such traces.

  \judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
              {\(\mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2\)}
         {\(\mpstatename \hookrightarrow \MPT_1 | f\)}

@@

\judgment{}{\(\obsT \lesssim \obsT\)}

\judgment{}{\(\tau \lesssim \obsT\)}

\judgmenttwo{\(\obsT_1 = \tau \obsT_1'\)}{\(\obsT_1' \lesssim \obsT_2\)}
            {\(\obsT_1 \lesssim \obsT_2\)}

\judgmenttwo{\(\obsT_2 = \tau \obsT_2'\)}{\(\obsT_1 \lesssim \obsT_2'\)}
            {\(\obsT_1 \lesssim \obsT_2\)}

\judgmentthree{\(\obsT_1 = w \obsT_1'\)}
              {\(\obsT_2 = w \obsT_2'\)}
              {\(\obsT_1' \lesssim \obsT_2'\)}
              {\(\obsT_1 \lesssim \obsT_2\)}

@@

% Fixed by Sean
%\leo{Why $\mathit{Obs}$ and not $\pi_o$ defined earlier?}
\judgmenttwo{\(\neg \isinf(\MPT)\)}{\(\pi_o(\MPT) \lesssim \pi_o(\machT)\)}
            {\(\MPT \sim_P \machT\)}

\judgmenttwo{\(\isinf(\MPT)\)}{\(\pi_o(\MPT) \simeq \pi_o(\machT)\)}
            {\(\MPT \sim_P \machT\)}

@@

 \[\mathit{Cof}(\mach,\funid,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and for some } \funid, \codemap(\component) = \funid \\
    (\HIGHSEC,\HIGHINT) & \text{else if } \component \in \WORDS
                          \text{ and } \component \leq \mach(\SP) - n \\
    (\HIGHSEC,\LOWINT) & \text{else if } \component \in \WORDS
                         \text{ and } \component \geq \mach(\SP) \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

@@

  \leo{Think more about whether we vary mc or step mc}
  \sna{I've left it at step mc for now}

@@

    Some forms of security rely on protected control flow. In a classic
    example, a system call that launches a missile might be gated behind code
    that checks a password. A piece of malicious code can jump into the middle
    of the sequence and execute the system call without the appropriate password,
    and still obey integrity if the malicious code never returns to its caller.
    Integrity here protects data, not privileges. But such a jump does violate
    eager stack confidentiality if the attacker jumps beyond its own function's
    instructions without performing a designated call or returning to an
    appropriate site.

@@

    While these ideas are here presented in a stack-centric style, they could
    be extended to heap safety in a system extended with a heap. Heap safety,
    like stack safety, may be framed as a form of noninterference
    \cite{AzevedoDeAmorim17}. As in stack safety, the mapping from components
    to security labels would be dynamic, with any heap data unreachable from
    the stack frame in a given state treated as high confidentiality and high
    integrity. This would require updating the contour at each step.

@@

    Return integrity captures the expected behavior of marked returns with
    regard to their call points. We begin by defining a coinductive relation
    \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_1 \MPT_2\). Given a call map
    \(\callmap\) and return map \(\retmap\), \(\MPT\) is related to
    \(\MPT'\) if \(\MPT'\) is a suffix of \(\MPT\) whose head is the first
    unmatched return in \(\MPT\). So if \(\MPT\) starts with a return,
    it is related to itself:

  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
           {\(\retmap(\mach(\PCname))\)}
      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\ 
        \mpstate{\mach}{\pol}\)}
  \judgmenttwo{\(\callmap(\mach(\PCname)) = \bot\)}
              {\(\retmap(\mach(\PCname))\)}
      {\(\mathit{Unmatched}\ \callmap\ \retmap\ \mpstate{\mach}{\pol}\MPT\ 
              \mpstate{\mach}{\pol}\MPT\)}

    If \(\MPT\) starts with a non-call, non-return state, its unmatched
    return must come later:

  \judgment{\(\callmap(\mach(\PCname)) = \bot
              \hspace{.5cm} \neg \retmap(\mach(\PCname))

              \mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT'\)}
           {\(\mathit{Unmatched}\ \callmap\ \retmap\ 
            \mpstate{\mach}{\pol}\MPT\ \MPT'\)}

    And most intricately, if \(\MPT\) starts with a call, we must find the
    return that matches it, which will be the unmatched return once we step
    past the call. Then the first unmatched return beyond it is the true
    unmatched return.

    \judgment{\(\callmap(\mach(\PCname)) = n
              \hspace{.5cm} \neg \retmap(\mach(\PCname))\)

              \(\MPT = \mpstate{\mach}{\pol} \MPT_{\mathit{call}}\)

              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{call}}\ \MPT_{\mathit{matched}}\)

              \(\MPT_{\mathit{matched}} = \mpstate{\mach'}{\pol'} \MPT_{\mathit{tail}}\)

              \(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT_{\mathit{tail}}\ \MPT_{\mathit{unmatched}}\)}
            {\(\mathit{Unmatched}\ \callmap\ \retmap\ \MPT\ \MPT_{\mathit{unmatched}}\)}
